The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at the European Southern Observatory, Paranal, Chile, Program ID: 088.A-0051 and 091.D-0904.}

### 2017-06-01 GRB111117A.tex
Nials comments
Closes #3

% TK: I made the following typographical changes:
% All physical constants are in italics, all mathematical constants are normal, so SFR
% used in formulas is italics, while it is not italics when used in text as an acronym
% Super- and subscripts are usually normal
% Write out all numbers < 12
% Added fixed spaces in GRB, e.g., GRB~111117A and units (this prevents a line break)
% Thin spaces \, between thousands when > 9999
% Table is never abbreviated
% Tried to make the text a bit easier to read, e.g., things like the following:
% Conclusions: Part of the reason is attributed ... -> The reason is ...

\abstract{
% TK I rephased the abstract a bit, check if you like it or not. The original version is commented out
It is notoriously difficult to localize short $\gamma$-ray bursts (sGRBs) and their hosts to measure their redshifts. These measurements, however, are critical to constrain sGRB progenitors 
and delay time models. Here, 
%Short $\gamma$-ray bursts (sGRBs) and their hosts are notoriously difficult to
%localize. 
we present spectroscopy of the host galaxy of GRB~111117A 
and measure its redshift to be $z = 2.211$. This makes GRB~111117A the most distant
high-confidence short GRB detected to date.
%with a secure redshift measurement. 
Our spectroscopic redshift supersedes a lower redshift previously estimated from photometry. %Spectroscopic observations along with a wealth of imaging data allow us to place tight
%constraints on the nature of the host galaxy at an unprecedented distance.
We use the spectroscopic redshift, as well as new imaging data to constrain the nature of the host galaxy as well as physical parameters of the GRB. The rest frame X-ray derived hydrogen column density, for example, is high compared to a complete sample of sGRBs and seems to follow the evolution with redshift as traced by the hosts of long GRBs (lGRBs). This could indicate that the part of the sGRB population hosted in late-type galaxies live in similar environments to
that of lGRB hosts.

The host lies in the brighter end of the expected host brightness distribution
at $z = 2.211$, and is actively forming stars. Using the host as a benchmark for
redshift determination, we assess that only 30 per cent of sGRB redshifts should
be missed due to host faintness at $z\sim2$. This indicates that we are not
missing redshifts for a dominant fraction of the sGRB hosts in this redshift range.
The high redshift of GRB~111117A is evidence against a lognormal delay-time model for
sGRBs through the predicted redshift distribution of sGRBs, which is very
sensitive to high-$z$ sGRBs.

@@

There is now mounting evidence that most short-duration gamma-ray bursts (GRBs)
come from the merger of neutron stars (NSs), either with another NS, or a black
hole, due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b}
supports this idea and distinguishes the physical origin of sGRBs from their
long-duration counterparts. %\citep[albeit see also][]{Fynbo2006b, Valle2006,
% Gal-Yam2006}.

% TK I rephrased this paragraph, i found the sentences too long and split . Original
% version is commented out.

The host galaxies of short GRBs are diverse. They are more massive and less actively 
star-forming on average than long GRB hosts \citep{Fong2013b}, while in some cases, no host 
galaxy can be identified. Together with their position 
within their hosts \citep{Fong2013a}, this suggests a progenitor system that can be very
long-lived, and is associated with stellar mass rather than star-formation rate
\citep{Berger2014}. The median redshift for sGRBs is $z\sim0.5$ \citep{Berger2014}, and
because most of these measurements come from the associated hosts, it is arguably biased
towards lower redshifts.

%The range of host galaxy types, in some cases apparently absent, and typically
%more massive and on average less actively star-forming than long GRB hosts 
%\citep{Fong2013b}, as well as their positions within hosts \citep{Fong2013a},
%suggest an origin for the progenitors that can be very long-lived, and
%associated with stellar mass rather than purely star-formation rate
%\citep{Berger2014}. The bulk of sGRBs have associated galaxies at relatively low
%redshifts with a median $z\sim0.5$ \citep{Berger2014} and because most redshifts
%are measured from the hosts, this value is biased towards low $z$.

%This progress on the progenitors of sGRBs has been enabled by the afterglow
% detection and burst localization by the \emph{Swift} satellite
% \citep{Gehrels2009a}. The bulk of these localizations have associated galaxies
% at relatively low redshifts with a median $z\sim0.5$ \citep{Berger2014} and
% because most redshifts are measured from the hosts, this value is biased
% towards low $z$.

@@

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, in four seperate epochs. The
burst was followed up 38 hours after the Burst Alert Telescope (BAT) trigger
under ESO programme 088.A-0051 (PI: Fynbo) and again later using a different ESO
program 091.D-0904 (PI: Hjorth). X-shooter can cover the wavelength range from
3000~\AA~to 24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across
three spectroscopic arms, simultaneously illuminated through the use of
dichroics. The bias-correction, flat-fielding, order tracing, wavelength
calibration, rectification, and flux calibration is then carried out using the
VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010}. 
%Because the spectra are curved across each detector, a rectification algorithm
%is employed which introduces correlations between neighboring pixels. We select
%a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the
%degree of correlation while conserving the maximal resolution.
The observations are combined and extracted using scripts described in Selsing
et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. The
signal-to-noise of the continuum in the near-infrared arm is too low to use the
optimal extraction algorithm \citep{Horne1986} and therefore the extraction is
carried out with a simple aperture.

@@

% TK: Some rephrasing below.

@@

% TK: I shortened a bit here
All data were reduced, analyzed and fitted in a similar manner as described in detail in
\citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}. Very briefly, we use our own
\texttt{python} and IRAF routines to perform a standard reduction which includes
bias/flat-field correction, de-fringing (if necessary), sky-subtraction, and
stacking of individual images. The photometry of the host was tied against magnitudes 
of field stars from the SDSS and 2MASS catalogs in the case of $grizJHK_{\mathrm{s}}$ filters. 
For the $R$ and $I$-band photometry we used the color transformations of
Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
We convert all magnitudes into the AB system if necessary, and correct for the Galactic 
foreground of $E_{B-V}=0.027~\mathrm{mag}$.

@@

We retrieved the automated data products provided by the \textit{Swift}-XRT GRB
repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}} \citep{Evans2009}. 
% TK: Fixed the link
The X-ray afterglow light curve
can be fit with a single power-law decay with index
$\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral analysis
using data in photon counting (PC) mode in the widest time epoch where the
$0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
data) to prevent spectral changes that can affect the X-ray column
density determination. The obtained spectrum is well described by an absorbed
power-law model. The best-fit spectral parameters are a photon index of $2.1 \pm
0.4$ and an intrinsic hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4} \times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a Galactic $N_{\mathrm{H}}$ in the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$.

@@

As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A securely 
belongs the short class of GRBs. Because the observed classification
indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift
\citep{Littlejohns2013a}, the updated redshift does not change this designation
significantly. The intrinsic luminosity increases, as visible in the X-ray light curve
in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
the majority of long GRBs. \citet{Bromberg2013} investigated the degree to which the
long and short population distributions overlap and quantified the certainty in
the class membership. GRB~111117A has $96_{-5}^{+3}$ percent probability of
being short. Compared to the other two short GRB candidates at high redshift, GRB060121
\citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
($17_{-15}^{+14}$ per cent) and GRB090426 \citep{Antonelli2009, Levesque2010,
	Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$ per cent), the certainty in class
membership for GRB~111117A is much higher.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Restframe $N_\mathrm{H}$}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We plot the recalculated $N_\mathrm{H}$ in Fig. \ref{fig:NH_z} where we compare with the
distributions of complete samples of both long and short GRBs. The long sample
is from \citet{Arcodia2016} and the short sample is from \citet{DAvanzo2014a}.
17 of the 99 long bursts do not have redshifts and likewise 5 out of 16 for
the short sample. Bursts without redshifts have been excluded for both groups.
GRB~111117A occupies a unique position in Fig. \ref{fig:NH_z} with the highest
$N_H$ of all short burst. The short sample, excluding GRB~111117A, is located at
low redshifts ($z < 1$) and is found to populate a similar column density
environment to long GRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
hydrogen column for GRB~111117A seems to follow the trend with increasing $N_H$
as a function of redshift as found for the long GRB afterglows \citep{Arcodia2016}.

@@

%From the clear host association, GRB~111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits clear emission lines indicative of a population of relatively young stars, 
As the majority of short GRBs \citep{Fong2013b}, the host of GRB~111117A is a
late-type galaxy and is entirely consistent in terms of stellar mass and stellar
age with the general population of short GRB hosts ($\left\langle M _*
\right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
0.3 $~Gyr) \citet{Leibler2010}. The SFR, on the other hand, is $\sim$1 order of
magnitude higher than the typical SFR for short GRB hosts galaxies
\citep{Berger2014} and more similar to the SFRs found in the hosts of long GRBs
at a corresponding redshift \citep{Kruhler2015}. Only two hosts in the sample of
short GRBs compiled in \citet{Berger2014} have a more vigorous star formation,
meaning that it is in the upper end of the star formation distribution,
while not a complete outlier. This is likely a selection effect, because a less
star-forming galaxy would exhibit weaker emission lines, thus making the
redshift harder to determine. Being a late-type host, both the stellar mass and
sSFR are entirely consistent with the expected host of
sGRBs \citep{Behroozi2014}.

The detection of \lya~is consistent with the SED-inferred absence of dust, despite
the moderate stellar age and the high X-ray derived hydrogen column density
which would suggest the opposite. The centroid of the Ly$\alpha$ emission is
found to be redshifted by $\sim$~220~km~s$^{-1}$ with respect to systemic, which
is similar to what is found for long GRB hosts \citep{Milvang-Jensen2012a} where
the outflow is attributed to star formation.

@@

A single sGRB at high redshift does little in terms of constraints on the
redshift distribution and therethrough the progenitor channels, but other short
hosts could be missed because they are intrinsically fainter and thus this
high-$z$ event is only detected due to the brightness of the host.
\citet{Berger2014} compiled a sample of sGRB host luminosities, normalized by
the characteristic galaxy luminosity at their respective redshift,
$L_B/L^{\star}_{B}$. 26 out of 39 hosts (66 per cent) in the sample have
redshifts. To convert the SED-inferred $M_B$ of GRB~111117A to
$L_B/L^{\star}_{B}$, we use the characteristic absolute B-band magnitude of the
Schechter function for the blue galaxies ($U - V < 0.25$) in the redshift window
$2.0 \leq z \leq 2.5$ from \citet{Marchesini2007} and find $L_B/L^{\star}_{B} =
1.2$, which is brighter than 70 per cent of the hosts in \citet{Berger2014} with
measured $L_B/L^{\star}_{B}$. If we assume that we are able to obtain emission-line
redshifts from hosts with $R < 25$~mag \citep{Kruhler2012}, then we would have
missed around 30\% (8 out of 26 from the sample of \citealt{Berger2014} with
measured $L_B/L^{\star}_{B}$), if they were at the redshift of GRB~111117A. If
we assume that the hosts missing redshift in \citet{Berger2014} follow the
redshift distribution of the hosts \textit{with} redshift, this implies that we
are \textit{not} missing a dominant fraction of sGRB redshift at $z \approx 2$,
due to host faintness.

@@

At $z = 2.211$, the age of the universe is 3 Gyr. If the progenitor systems
of sGRBs are the merger of two neutron stars, this sets a hard upper limit to
the coalescence timescale for such a system. In the absence of other mechanisms,
the timescale of the orbital decay of the system is set by the energy loss due
to gravitational waves, which in turn is set by the mass of constituent compact
objects and the separation of the two \citep{Postnov2014}. If we assume that the
formation timescale of the first galaxies is short compared to the time since
the big bang \citep{Richard2011} and if we assume a mass of 1.4~$M_\odot$ for
each of the neutron stars at the time of system formation, this places a limit
on the initial separation of the two neutron stars of $a_0 < 3.2~R_\odot$,
where $a_0$ is the initial separation.

Assuming our measured age of the stellar population ($\tau = 0.5$~Gyr) as a
limit for the orbital decay time we can put a tighter constraint of $a_0 < 2.1~R_\odot$. 
However, this does not account for the possibility there could be an
underlying stellar population of older stars from a previous star-formation
epoch, which is why we consider $a_0 < 3.2~R_\odot$ as a hard upper limit. The
delay time between formation and explosion is well accommodated by the models of
\citet{Belczynski2006}, although the longest formation channels are excluded.
This is especially true given the late type nature of the host
\citep{OShaughnessy2008}.

@@

In this \emph{Letter}, we have provided a revised, spectroscopic redshift for the short
GRB~111117A based on emission lines %discovered using an improved reduction scheme, 
setting it at $z = 2.211$. This value supersedes and corrects the previous
% TK: No need to repeat a wrong redshift here, so slightly rephrased
photometric redshift of \citet{Margutti2012} and \citet{Sakamoto2013}. The erroneous redshift estimate of previous authors is attributed to a discrepancy in the measured $z$-band magnitude.

@@

Although a single burst carries little leverage in terms of constraining the
redshift distribution of sGRB, the high redshift of GRB~111117A needs to be
accommodated in progenitor models. A lognormal delay time model predicts a very low volumetric
density of bursts at $z = 2.211$, whereas a power law delay time model peaks
near GRB~111117A. If more sGRBs are at similarly high redshifts, but are missed due to
the faintness of their hosts, a lognormal delay time model will be disfavored.
Compared to a sample of short GRB hosts, GRB~111117A is more luminous than 70 per
cent of the sample with measured luminosities. For 30 per
cent of the host sample, we would be unable to determine a redshift should they be at 
a similar redshift as GRB~111117A. This implies that we are \textit{not} missing a dominant
fraction of the sGRBs at $z \sim 2$.

>>>>

\abstract{
It is notoriously difficult to localize short $\gamma$-ray bursts (sGRBs) and
their hosts to measure their redshifts. These measurements, however, are
critical to constrain sGRB progenitors and delay time models. Here, we present
spectroscopy of the host galaxy of GRB~111117A and measure its redshift to be $z
= 2.211$. This makes GRB~111117A the most distant high-confidence short GRB
detected to date. Our spectroscopic redshift supersedes a lower redshift value
for this burst previously estimated from photometry.

We use the spectroscopic redshift, as well as new imaging data to constrain the
nature of the host galaxy as well as physical parameters of the GRB. The rest
frame X-ray derived hydrogen column density, for example, is high compared to a
complete sample of sGRBs and seems to follow the evolution with redshift as
traced by the hosts of long GRBs (lGRBs). This is consistent with the proportion
of the sGRB population originating in late-type galaxies live in similar
environments to that of lGRB hosts.

The host lies in the brighter end of the expected host brightness distribution
at $z = 2.211$, and is actively forming stars. Using the host as a benchmark for
redshift determination, we find that less than 55 per cent of all sGRB redshifts
should be missed due to host faintness at $z\sim2$. The high redshift of
GRB~111117A is evidence against a lognormal delay-time model for sGRBs through
the predicted redshift distribution of sGRBs, which is very sensitive to
high-$z$ sGRBs.

@@

There is now mounting evidence that most short-duration gamma-ray bursts (GRBs)
come from the merger of neutron stars (NSs), either with another NS, or a black
hole, due to their apparent association with kilonovae \citep{Barnes2013a,
	Tanvir2013b, Yang2015, Jin2016, Rosswog2016}. The absence of associated
supernovae in deep searches \citep[e.g.][]{Hjorth2005a,Fox2005,Hjorth2005b}
supports this idea and distinguishes the physical origin of sGRBs from their
long-duration counterparts. %\citep[albeit see also][]{Fynbo2006b, Valle2006, Gal-Yam2006}.

The host galaxies of short GRBs are diverse. They are more massive and less actively 
star-forming on average than long GRB hosts \citep{Fong2013b}, while in some cases, no host 
galaxy can be identified \citep{Berger2010a, Tunnicliffe2014}. Together with their position 
within their hosts \citep{Fong2013a}, this suggests a progenitor system that can be very
long-lived, and is associated with stellar mass rather than star-formation rate
\citep{Berger2014}. The median redshift for sGRBs is $z\sim0.5$ \citep{Berger2014}, and
because most of these measurements come from the associated hosts, it is arguably biased
towards lower redshifts.

@@

Spectroscopic observations were carried out using the cross-dispersed echelle
spectrograph, VLT/X-shooter \citep{Vernet2011}, at four seperate epochs. The
burst was followed up 38 hours after the Burst Alert Telescope (BAT) trigger
under ESO programme 088.A-0051 (PI: Fynbo) and again later using a different ESO
program 091.D-0904 (PI: Hjorth). X-shooter can cover the wavelength range from
3000~\AA~to 24\,800~\AA~(21\,000~\AA~when the $K$-band blocking filter is used) across
three spectroscopic arms, simultaneously illuminated through the use of
dichroics. The bias-correction, flat-fielding, order tracing, wavelength
calibration, rectification, and flux calibration is then carried out using the
VLT/X-shooter pipeline version 2.8.4 \citep{Modigliani2010}. 
%Because the spectra are curved across each detector, a rectification algorithm
%is employed which introduces correlations between neighboring pixels. We select
%a pixel-scale of 0.2/0.2/0.6 \AA/pix for the UVB/VIS/NIR arm to minimize the
%degree of correlation while conserving the maximal resolution.
The observations are combined and extracted using scripts described in Selsing
et al. 2017 (in prep.) and available
online\footnote{\url{https://github.com/jselsing/XSGRB_reduction_scripts}}. The
signal-to-noise of the continuum in the near-infrared arm is too low to use the
optimal extraction algorithm \citep{Horne1986} and therefore the extraction is
carried out with a simple aperture.

@@

@@

All data were reduced, analyzed and fitted in a similar manner as described in detail in
\citet{Kruhler2011a} and, more recently, in \citet{Schulze2016}. Very briefly, we use our own
\texttt{python} and IRAF routines to perform a standard reduction which includes
bias/flat-field correction, de-fringing (if necessary), sky-subtraction, and
stacking of individual images. The photometry of the host was tied against magnitudes 
of field stars from the SDSS and 2MASS catalogs in the case of $grizJHK_{\mathrm{s}}$ filters. 
For the $R$ and $I$-band photometry we used the color transformations of
Lupton\footnote{\url{https://www.sdss3.org/dr8/algorithms/sdssUBVRITransform.php}}. 
We convert all magnitudes into the AB system if necessary, and correct for the Galactic 
foreground of $E_{B-V}=0.027~\mathrm{mag}$.

@@

We retrieved the automated data products provided by the \textit{Swift}-XRT GRB
repository\footnote{\url{http://www.swift.ac.uk/xrt\_products/00507901}}
\citep{Evans2009}. The X-ray afterglow light curve can be fit with a single power-law decay with
index $\alpha=1.27_{-0.10}^{+0.12}$. We performed a time-integrated spectral
analysis using data in photon counting (PC) mode in the widest time epoch where
the $0.3-1.5\,\mathrm{keV}$ to $1.5-10\,\mathrm{keV}$ hardness ratio is constant
(namely, from $t-T_0 = 205$~s to $t-T_0 = 203.5$~ks, for a total of 29.1~ks of
data) to prevent spectral changes that can affect the X-ray column density
determination. The obtained spectrum is well described by an absorbed power-law
model. The best-fit spectral parameters are a photon index of $2.1 \pm 0.4$ and
an intrinsic hydrogen column density $N_{\mathrm{H}}$ of $2.4_{-1.6}^{+2.4}
\times 10^{22}$~cm$^{-2}$ ($z=2.211$), assuming a Galactic $N_{\mathrm{H}}$ in
the burst direction of $4.1 \times 10^{20}$~cm$^{-2}$ \cite{Willingale2013}.

@@

As already pointed out \citep{Margutti2012, Sakamoto2013}, GRB~111117A securely 
belongs the short class of GRBs. Because the observed classification
indicators, $T_{90}$ and hardness ratio, do not depend strongly on redshift
\citep{Littlejohns2013a}, the updated redshift does not change this designation
significantly. The intrinsic luminosity increases, as visible in the X-ray light curve
in Fig. \ref{fig:sxray_lightcurve}, but it still is sub-luminous compared to
the majority of long GRBs. \citet{Bromberg2013} investigated the degree to which the
long and short population distributions overlap and quantified the certainty in
the class membership. GRB~111117A has $96_{-5}^{+3}$ percent probability of
being short. Compared to the other two sGRB candidates at high redshift, GRB060121
\citep{DeUgartePostigo2006, Levan2006} at $1.7 \lesssim z \lesssim 4.5$
($17_{-15}^{+14}$ per cent) and GRB090426 \citep{Antonelli2009, Levesque2010,
	Thone2011} at $z = 2.609$ ($10_{-10}^{+15}$ per cent), the certainty in class
membership for GRB~111117A is much higher.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Restframe $N_\mathrm{H}$} \label{restnH}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We plot the recalculated $N_\mathrm{H}$ in Fig. \ref{fig:NH_z} where we compare with the
distributions of complete samples of both long and short GRBs. The lGRB sample
is from \citet{Arcodia2016} and the sGRB sample is from \citet{DAvanzo2014a}.
17 of the 99 long bursts do not have redshifts and likewise 5 out of 16 for
the short sample. Bursts without redshifts have been excluded for both groups.
GRB~111117A occupies a unique position in Fig. \ref{fig:NH_z} with the highest
$N_H$ of all short burst. The short sample, excluding GRB~111117A, is located at
low redshifts ($z < 1$) and is found to populate a similar column density
environment to long GRBs at similar redshifts \citep{DAvanzo2014a}. The inferred
hydrogen column for GRB~111117A seems to follow the trend with increasing $N_H$
as a function of redshift as found for the long GRB afterglows \citep{Arcodia2016}.

@@

%From the clear host association, GRB~111117A does not belong to the hostless class of GRBs \citep{Berger2010a} and because the host exhibits emission lines indicative of a population of relatively young stars, 
As the majority of short GRBs \citep{Fong2013b}, the host of GRB~111117A is a
late-type galaxy and is entirely consistent in terms of stellar mass and stellar
age with the general population of short GRB hosts ($\left\langle M _*
\right\rangle = 10^{10.1} M_{\odot}$ and $\left\langle \tau _* \right\rangle =
0.3 $~Gyr) \citep{Leibler2010}. Being a late-type host, both the stellar mass
and sSFR are entirely within the range expected for the hosts of sGRBs
\citep{Behroozi2014}. The SFR is $\sim$1 order of magnitude higher than the
typical SFR for short GRB hosts galaxies \citep{Berger2014} and more similar to
the SFRs found in the hosts of long GRBs at a corresponding redshift
\citep{Kruhler2015}. The high SFR is partly a selection effect, because a less
star-forming galaxy would exhibit weaker emission lines, thus making the
redshift harder to determine. Additionally, it is natural to expect some
evolution in the hosts of sGRBs with redshift as illustrated in Sect.
\ref{restnH}.

%The detection of \lya~is consistent with the SED-inferred absence of dust, despite
%the moderate stellar age and the high X-ray derived hydrogen column density
%which would suggest the opposite. The centroid of the Ly$\alpha$ emission is
%found to be redshifted by $\sim$~220~km~s$^{-1}$ with respect to systemic, which
%is similar to what is found for long GRB hosts \citep{Milvang-Jensen2012a} where
%the outflow is attributed to star formation.

@@

A single sGRB at high redshift does little in terms of constraining the redshift
distribution and therethrough the progenitor channels. In particular, other
short-GRB hosts could be missed because they are intrinsically fainter and thus
this high-$z$ event is only detected due to the brightness of the host.
\citet{Berger2014} compiled a sample of sGRB host luminosities, normalized by
the characteristic galaxy luminosity at their respective redshift,
$L_B/L^{\star}_{B}$. 26 out of 39 hosts (66 per cent) in the sample have
redshifts. To convert the SED-inferred $M_B$ of GRB~111117A to
$L_B/L^{\star}_{B}$, we use the characteristic absolute B-band magnitude of the
Schechter function for the blue galaxies ($U - V < 0.25$) in the redshift window
$2.0 \leq z \leq 2.5$ from \citet{Marchesini2007} and find $L_B/L^{\star}_{B} =
1.2$, which is brighter than 70 per cent of the hosts in \citet{Berger2014} with
measured $L_B/L^{\star}_{B}$.

If we assume that we are able to obtain emission-line redshifts from hosts with
$R < 25$~mag \citep{Kruhler2012}, then we would have missed around 30 per cent
(8 out of 26 from the sample of \citealt{Berger2014} with measured
$L_B/L^{\star}_{B}$), if they were at the redshift of GRB~111117A. Because the
average SFR of galaxies hosting lGRBs is higher than for galaxies hosting sGRBs,
the fraction of missed burst redshifts is likely higher although the cosmic SFR
evolution could play a role in improving redshift determinability.

%If we additionally assume that the hosts missing redshift in
%\citet{Berger2014} follow the redshift distribution of the hosts \textit{with}
%redshift, this implies that we are \textit{not} missing a dominant fraction of
%sGRB redshift at $z \approx 2$, due to host faintness.

A fraction of the bursts missing redshift are host-less and is therefore likely
at moderate redshifts \citep{Tunnicliffe2014}, but should some of the remainder
be at high redshift, the missed fraction will increase. If we assume that
\textit{all} the bursts that are missing redshifts in \citet{Berger2014} are at
high-$z$ and missed due to host faintness, 22 out of 39 hosts (55 per cent)
would be missed at $z = 2.211$. This serves as an upper limit on the fraction of
missed burst at high-$z$ and illustrates that we are likely \textit{not} missing a
large fraction of sGRBs redshift at $z \approx 2$, due to host faintness.

@@

At $z = 2.211$, the age of the universe is 3 Gyr. If the progenitor systems of
sGRBs are the merger of two neutron stars, this sets a hard upper limit to the
coalescence timescale for such a system. In the absence of other mechanisms, the
timescale of the orbital decay of the system is set by the energy loss due to
gravitational waves, which in turn is set by the mass of constituent compact
objects and the separation of the two \citep{Postnov2014}. If we assume that the
formation timescale of the first galaxies is short compared to the time since
the big bang \citep{Richard2011} and if we assume a mass of 1.4~$M_\odot$ for
each of the neutron stars at the time of system formation, this places a hard
upper limit on the initial separation, $a_0$, of the two neutron stars of $a_0 <
3.2~R_\odot$.

If we use the stellar population age from our SED fit, then we obtain a
(softer) limit on the initial separation of $a_0 < 2.1~R_\odot$. However, this
does not account for the possibility there could be an underlying stellar
population of older stars from a previous star-formation epoch. The delay time
between formation and explosion is well accommodated by the models of
\citet{Belczynski2006}, although the longest formation channels are excluded.
This is especially true given the late type nature of the host
\citep{OShaughnessy2008}.

@@

In this \emph{Letter}, we have provided a revised, spectroscopic redshift for the short
GRB~111117A based on emission lines %discovered using an improved reduction scheme, 
setting it at $z = 2.211$. This value supersedes and corrects the previous
photometric redshift of \citet{Margutti2012} and \citet{Sakamoto2013}. The erroneous redshift estimate of previous authors is attributed to a discrepancy in the measured $z$-band magnitude.

@@

Although a single burst carries little leverage in terms of constraining the
redshift distribution of sGRB, the high redshift of GRB~111117A needs to be
accommodated in progenitor models. A lognormal delay time model predicts a very low volumetric
density of bursts at $z = 2.211$, whereas a power law delay time model peaks
near GRB~111117A. If more sGRBs are at similarly high redshifts, but are missed due to
the faintness of their hosts, a lognormal delay time model will be disfavored.
Compared to a sample of short GRB hosts, GRB~111117A is more luminous than 70 per
cent of the sample with measured luminosities. Conservatively, for 55 per
cent of the sGRB hosts, we would be unable to determine a redshift should they be at 
a similar redshift as GRB~111117A. This implies that we are \textit{not} missing a large
fraction of the sGRBs at $z \sim 2$.
